Regression-based

### 2023-12-16 Final Submission.Rmd
SUBMISSION is our final file

---
title: "Final Project Report: An exploration of Biliary Cirrhosis and Treatments"
output: html_notebook
authors: "Mihir Kulkarni, Nithika Menon"
---

@@

@@

![[Primary Biliary Cholangitis ([Source](https://www.google.com/search?client=safari&sca_esv=591276386&rls=en&sxsrf=AM9HkKmtJVOC2Jrf2BSYVB2O-QcnTSgx6g:1702664705100&q=Primary+Biliary+Cholangitis&tbm=isch&source=lnms&sa=X&ved=2ahUKEwid0qyniJKDAxX9FFkFHVH3CcoQ0pQJegQIDxAB&biw=1440&bih=848&dpr=2#imgrc=UeV_fhP-8vPOlM))]{.underline}](17715-primary-biliary-cholangitis.webp)

@@

As, you can see, the model was able to create clear clusters for the data. However, it is important to find the best value of k to get the optimal clusters. To do this we can use a scree plot. Here, we\'re looking for the number of clusters that corresponds to the \"elbow\".

@@

@@

>>>>

---
title: 'Final Project Report - A statistical exploration of Biliary Cirrhosis and Treatments'
output:
  pdf_document: default
authors: Mihir Kulkarni+Nithika Menon
---

```{r setupFiles, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)
```

@@

![[Primary biliary cholangitis (Source:]{.underline} <https://www.mayoclinic.org/diseases-conditions/primary-biliary-cholangitis/symptoms-causes/syc-20376874>)](ds00604_-ds00918_im02271_r7_biliarysystemthu_jpg.jpg)

@@

@@

As, you can see, the model was able to create clear clusters for the data. However, it is important to find the best value of k to get the optimal clusters. To do this we can use a scree plot. Here, we're looking for the number of clusters that corresponds to the "elbow".

@@

\newpage

@@

```{r codeAppendix}
#| ref.label = knitr::all_labels(),
#| echo = TRUE,
#| eval = FALSE
```

### 2023-12-16 Final-Submission.tex
SUBMISSION is our final file

  pdftitle={Submission},

@@

\title{Submission}
\author{}
\date{\vspace{-2.5em}}

@@

{
\setcounter{tocdepth}{2}
\tableofcontents
}
\hypertarget{background}{%
\section{Background}\label{background}}

@@

@@

\hypertarget{q1-what-predictors-if-any-increases-an-individuals-chance-of-receiving-a-liver-transplant}{%
\subsubsection{Q1: What predictors, if any, increases an individual's
chance of receiving a liver
transplant?}\label{q1-what-predictors-if-any-increases-an-individuals-chance-of-receiving-a-liver-transplant}}

\hypertarget{q2-are-there-factors-not-currently-included-in-medical-definition-that-can-help-distinguish-different-stages-of-biliary-cholangitis}{%
\subsubsection{Q2: Are there factors not currently included in medical
definition that can help distinguish different stages of Biliary
Cholangitis?}\label{q2-are-there-factors-not-currently-included-in-medical-definition-that-can-help-distinguish-different-stages-of-biliary-cholangitis}}

@@

The dataset is produced from a clinical study by Mayo Clinic run from
1974-1984. The final `Status' of each patient was observed in 1986.
`Status' was either Dead, Censored, or Censored due to liver transplant.

A link to the original study is here:
\url{https://faculty.washington.edu/abansal/ShortCourse_DynamicDecisionMaking/Dickson1989_MayoPBCOriginalArticle.pdf}

\hypertarget{load-in-packages}{%
\subsubsection{Load in Packages}\label{load-in-packages}}

@@

\hypertarget{data-processing}{%
\subsubsection{Data Processing}\label{data-processing}}

@@

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{head}\NormalTok{(cirrhosis)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # A tibble: 6 x 20
##      ID N_Days Status Drug          Age Sex   Ascites Hepatomegaly Spiders Edema
##   <dbl>  <dbl> <chr>  <chr>       <dbl> <chr> <chr>   <chr>        <chr>   <chr>
## 1     1    400 D      D-penicill~ 21464 F     Y       Y            Y       Y    
## 2     2   4500 C      D-penicill~ 20617 F     N       Y            Y       N    
## 3     3   1012 D      D-penicill~ 25594 M     N       N            N       S    
## 4     4   1925 D      D-penicill~ 19994 F     N       Y            Y       S    
## 5     5   1504 CL     Placebo     13918 F     N       Y            Y       N    
## 6     6   2503 D      Placebo     24201 F     N       Y            N       N    
## # i 10 more variables: Bilirubin <dbl>, Cholesterol <dbl>, Albumin <dbl>,
## #   Copper <dbl>, Alk_Phos <dbl>, SGOT <dbl>, Tryglicerides <dbl>,
## #   Platelets <dbl>, Prothrombin <dbl>, Stage <dbl>
\end{verbatim}

\hypertarget{pre-process-the-data}{%
\subsubsection{Pre-process the data}\label{pre-process-the-data}}

@@

\begin{itemize}
\item
  Engage in Exploratory Data Analysis
\item
  Use Machine Learning Algorithms:

  \begin{itemize}
  \tightlist
  \item
    You must use at least one of each: a Regression-based algorithm, a
    Tree-based algorithm, and an Unsupervised Learning algorithm.
  \item
    Each algorithm must be explained for a general audience.
  \item
    All results should be explained.
  \item
    Where appropriate, comparisons between methods should be discussed.
    This includes highlighting the strengths and weaknesses of the
    various approaches as well as when each is appropriate.
  \end{itemize}
\item
  A discussion of what you've learned from your work about the context
  and recommendations for future work.
\item
  Author Contributions--list out who did what in the project.
\item
  References--Citation style is your choice, but all sources should be
  documented (both in text and in the References section). This includes
  where you got your data.
\item
  A commented Code Appendix
\end{itemize}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{cirrhosis }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ N\_Days)) }\SpecialCharTok{+} \FunctionTok{geom\_histogram}\NormalTok{() }\SpecialCharTok{+}
  \FunctionTok{facet\_wrap}\NormalTok{(}\AttributeTok{facets =} \FunctionTok{vars}\NormalTok{(Status))}
\end{Highlighting}
\end{Shaded}

@@

\includegraphics{Submission_files/figure-latex/unnamed-chunk-9-1.pdf}

@@

\hypertarget{eda}{%
\section{EDA}\label{eda}}

\hypertarget{exploring-the-demographics-of-the-patients}{%
\subsection{Exploring the demographics of the
patients}\label{exploring-the-demographics-of-the-patients}}

@@

\includegraphics{Submission_files/figure-latex/unnamed-chunk-11-1.pdf}

@@

\includegraphics{Submission_files/figure-latex/unnamed-chunk-12-1.pdf}

@@

\hypertarget{looking-at-the-data-itself}{%
\subsection{Looking at the data
itself}\label{looking-at-the-data-itself}}

@@

An observation is that many of the variables have 106 NAs. This
indicates that a good fraction {[}106 patients{]} were equally tracked
and measured in a less extensive way.

@@

\includegraphics{Submission_files/figure-latex/unnamed-chunk-14-1.pdf}

@@

\hypertarget{exploring-some-interactions-of-variables}{%
\subsection{Exploring some interactions of
variables}\label{exploring-some-interactions-of-variables}}

@@

\includegraphics{Submission_files/figure-latex/unnamed-chunk-15-1.pdf}

@@

\includegraphics{Submission_files/figure-latex/unnamed-chunk-16-1.pdf}

@@

\hypertarget{mihir-choose-your-preffered-graph-and-explain}{%
\subsubsection{mihir choose your preffered graph and
explain}\label{mihir-choose-your-preffered-graph-and-explain}}

@@

\includegraphics{Submission_files/figure-latex/unnamed-chunk-17-1.pdf}

@@

\includegraphics{Submission_files/figure-latex/unnamed-chunk-18-1.pdf}

@@

\hypertarget{part-i-classification-algorithm-for-the-status-of-the-patient}{%
\section{Part I: Classification Algorithm for the Status of the
Patient}\label{part-i-classification-algorithm-for-the-status-of-the-patient}}

@@

In this report, I have decided to use \{rpart\} because it provides more
flexibility for surrogate splits and the trees are a bit easier to make
attractive looking.

@@

\includegraphics{Submission_files/figure-latex/unnamed-chunk-22-1.pdf}

(Explanation of above) To gain a further understanding of the data, we
can plot a tree yielding Collection Node style trees. This can help us
understand how the data is split.

@@

\includegraphics{Submission_files/figure-latex/unnamed-chunk-23-1.pdf}

@@

\includegraphics{Submission_files/figure-latex/unnamed-chunk-25-1.pdf}

From the graph we can see that a cp of 0.011 is ideal as it is under the
horizontal (dotted) reference line. We can prune the tree with this CP
value.

@@

\includegraphics{Submission_files/figure-latex/unnamed-chunk-27-1.pdf}

@@

\FunctionTok{kable}\NormalTok{(}
\NormalTok{  conf\_matrix,}
  \AttributeTok{col.names =} \FunctionTok{c}\NormalTok{(}\StringTok{"Prediction/Supervision"}\NormalTok{, }\StringTok{"0"}\NormalTok{, }\StringTok{"1"}\NormalTok{),}
  \AttributeTok{digits =} \DecValTok{3}\NormalTok{,}
  \AttributeTok{booktabs =} \ConstantTok{TRUE}\NormalTok{,}
  \AttributeTok{caption =} \StringTok{"Model 1: Confusion Matrix"}\NormalTok{,}
  \AttributeTok{align =} \StringTok{"c"}
\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
\FunctionTok{kable\_styling}\NormalTok{(}\AttributeTok{latex\_options =} \StringTok{"HOLD\_position"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\caption{\label{tab:unnamed-chunk-29}Model 1: Confusion Matrix}
\centering
\begin{tabular}[t]{lcc}
\toprule
Prediction/Supervision & 0 & 1\\
\midrule
0 & 19 & 9\\
1 & 14 & 43\\
\bottomrule
\end{tabular}
\end{table}

@@

From this, we can get the model's

\begin{itemize}
\item
  True Positive (TP): Predicted Graduated and actually Graduated. In
  this case, 2217
\item
  False Positive (FP): Predicted Graduated but actually Not Graduated.
  In this case, 1000
\item
  False Negative (FN): Predicted Not Graduated but actually Graduated.
  In this case, 244
\item
  True Negative (TN): Predicted Not Graduated and actually Not
  Graduated. In this case, 659
\end{itemize}

With this, we can also calculate the model's metrics.

@@

@@

\caption{\label{tab:unnamed-chunk-36}Model 1 Confusion Matrix}
\centering
\begin{tabular}[t]{ccc}
\toprule
Predicted/Actual & Censored & Deceased\\
\midrule
Censored & 117 & 68\\
Deceased & 15 & 20\\
\bottomrule
\end{tabular}
\end{table}

@@

This analysis suggests significant associations between several
predictors (Bilirubin, Age, Alk\_Phos, Prothrombin, Spiders) and the
status of the patient. Each predictor, including the Intercept, shows a
statistically significant effect, as indicated by their small p-values
(Pr(\textgreater\textbar z\textbar)). The negative coefficients for
Bilirubin, Age, Alk\_Phos, Prothrombin, and Spiders imply that increases
in these variables are associated with a decrease in the log-odds of the
outcome. Bilirubin and Spiders are showing notably strong effects in the
model.

@@

\includegraphics{Submission_files/figure-latex/unnamed-chunk-38-1.pdf}

This figure shows us the Tukey-Anscombe plot using Pearson residuals for
Model 2. In an ideal fit, the residuals should be evenly distributed
about zero with constant mean and variance. The shape of the line
suggests that the model is not capturing someunder lying structure in
the data, especially in extreme cases.

@@

\caption{\label{tab:unnamed-chunk-40}Confusion matrix for Model 2}
\centering
\begin{tabular}[t]{l|r|r}
\hline
  & Censored & Deceased\\
\hline
Censored & 115 & 24\\
\hline
Deceased & 17 & 64\\
\hline
\end{tabular}
\end{table}

@@

\includegraphics{Submission_files/figure-latex/unnamed-chunk-42-1.pdf}

@@

\includegraphics{Submission_files/figure-latex/unnamed-chunk-43-1.pdf}

@@

\includegraphics{Submission_files/figure-latex/unnamed-chunk-45-1.pdf}

@@

Model 1: - Its ROC curve is above the line of no discrimination,
indicating that the model has some predictive capabilities. - The AUC is
0.64, which is better than random guessing but suggests there's room for
improvement.

Model 2: - The ROC curve for Model 2 is significantly above that of
Model 1, and much closer to the top-left corner, indicating better
predictive performance. - The AUC is 0.899, which suggests a good
classification performance, and it's notably better than Model 1. - Its
ability to discriminate between positive and negative classes is
superior to that of Model 1.

@@

\includegraphics{Submission_files/figure-latex/unnamed-chunk-46-1.pdf}

@@

\caption{\label{tab:unnamed-chunk-48}Confusion Matrix for Test data}
\centering
\begin{tabular}[t]{l|r|r}
\hline
Predicted/Actual & Censored & Deceased\\
\hline
Censored & 30 & 13\\
\hline
Deceased & 3 & 10\\
\hline
\end{tabular}
\end{table}

@@

\includegraphics{Submission_files/figure-latex/unnamed-chunk-49-1.pdf}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#do the ml, take about observational stages vs now quantify, look into the different variables and what they mean.}
\NormalTok{cirrhosisCluster }\OtherTok{\textless{}{-}}\NormalTok{ cirrhosis}

@@

\includegraphics{Submission_files/figure-latex/unnamed-chunk-54-1.pdf}

@@

\includegraphics{Submission_files/figure-latex/unnamed-chunk-55-1.pdf}

@@

\includegraphics{Submission_files/figure-latex/unnamed-chunk-57-1.pdf}

@@

\includegraphics{Submission_files/figure-latex/unnamed-chunk-58-1.pdf}

@@

\includegraphics{Submission_files/figure-latex/unnamed-chunk-59-1.pdf}

>>>>

  pdftitle={Regression-based},
  pdfauthor={Mihir Kulkarni, Nithika Menon},

@@

\title{Regression-based}
\author{Mihir Kulkarni, Nithika Menon}
\date{2023-12-12}

@@

\#Final Project Report-An exploration of Biliary Cirrhosis and
Treatments

\hypertarget{background}{%
\section{Background}\label{background}}

@@

\hypertarget{why-is-the-data-important-to-us}{%
\subsection{Why is the data important to
us?}\label{why-is-the-data-important-to-us}}

Both of us have been very interetsted in the intersection of medicine
and machine learning. The idea that the concepts we are learning in our
DS majors could have such an impact on the health industry was really
special. Another important aspect was that there is so much potential
for advancement of medicine, we both agreed that this area is MLs most
promising aspect. The cirrhosis data set perticularly stood out to us
because the data set included so many interesting attributes and we
anted to see how they are connected with the outcomes of the patient.

The data set also includes some important limitations. Most importantly,
there are only 418 cases, greatly affecting the types of analysis we
could do. This shows realm world problems data scientists face. Thus,
our goal is to use our Statistical DS skillset to approach this problem,
despite the limitations.

@@

Exploring the study, we concluded on two core questions we hope to
address in this report:

\hypertarget{crq1-what-predictors-if-any-affected-the-status-of-the-patient-at-the-end-of-the-study}{%
\subsubsection{CRQ1: What predictors, if any affected the status of the
patient at the end of the
study?}\label{crq1-what-predictors-if-any-affected-the-status-of-the-patient-at-the-end-of-the-study}}

\hypertarget{crq2-are-there-factors-not-currently-included-in-medical-definition-that-can-help-distinguish-different-stages-of-biliary-cholangitis}{%
\subsubsection{CRQ2: Are there factors not currently included in medical
definition that can help distinguish different stages of Biliary
Cholangitis?}\label{crq2-are-there-factors-not-currently-included-in-medical-definition-that-can-help-distinguish-different-stages-of-biliary-cholangitis}}

@@

The dataset is produced from a clinical study by Mayo Clinic run from
1974-1984. The final `Status' of each patient was observed in 1986.
`Status' was either Dead, Censored, or Censored due to liver transplant.
A link to the original study is here:
\url{https://faculty.washington.edu/abansal/ShortCourse_DynamicDecisionMaking/Dickson1989_MayoPBCOriginalArticle.pdf}

The data contains the following attributes
(\href{https://www.kaggle.com/datasets/fedesoriano/cirrhosis-prediction-dataset/data}{Data
Source}):

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  \textbf{ID}: unique identifier\\
\item
  \textbf{N\_Days}: number of days between registration and the earlier
  of death, transplantation, or study analysis time in July 1986\\
\item
  \textbf{Status}: status of the patient C (censored), CL (censored due
  to liver tx), or D (death)\\
\item
  \textbf{Drug}: type of drug D-penicillamine or placebo\\
\item
  \textbf{Age}: age in {[}days{]}\\
\item
  \textbf{Sex}: M (male) or F (female)\\
\item
  \textbf{Ascites}: presence of ascites N (No) or Y (Yes)\\
\item
  \textbf{Hepatomegaly}: presence of hepatomegaly N (No) or Y (Yes)\\
\item
  \textbf{Spiders}: presence of spiders N (No) or Y (Yes)\\
\item
  \textbf{Edema}: presence of edema N (no edema and no diuretic therapy
  for edema), S (edema present without diuretics, or edema resolved by
  diuretics), or Y (edema despite diuretic therapy)\\
\item
  \textbf{Bilirubin}: serum bilirubin in {[}mg/dl{]}\\
\item
  \textbf{Cholesterol}: serum cholesterol in {[}mg/dl{]}\\
\item
  \textbf{Albumin}: albumin in {[}gm/dl{]}\\
\item
  \textbf{Copper}: urine copper in {[}ug/day{]}\\
\item
  \textbf{Alk\_Phos}: alkaline phosphatase in {[}U/liter{]}\\
\item
  \textbf{SGOT}: SGOT in {[}U/ml{]}\\
\item
  \textbf{Triglycerides}: triglicerides in {[}mg/dl{]}\\
\item
  \textbf{Platelets}: platelets per cubic {[}ml/1000{]}\\
\item
  \textbf{Prothrombin}: prothrombin time in seconds {[}s{]}\\
\item
  \textbf{Stage}: histologic stage of disease (1, 2, 3, or 4)
\end{enumerate}

For the basic pre-processing, we binary-encoded the categorical
variables in the data set and set them as factors.

@@

@@

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#?????}
\NormalTok{cirrhosis }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ N\_Days)) }\SpecialCharTok{+} \FunctionTok{geom\_histogram}\NormalTok{() }\SpecialCharTok{+}
  \FunctionTok{facet\_wrap}\NormalTok{(}\AttributeTok{facets =} \FunctionTok{vars}\NormalTok{(Status))}
\end{Highlighting}
\end{Shaded}

@@

\includegraphics{Final-Submission_files/figure-latex/unnamed-chunk-3-1.pdf}

@@

\hypertarget{exploratory-data-analysis}{%
\section{Exploratory Data Analysis}\label{exploratory-data-analysis}}

\hypertarget{part-1-exploring-the-demographics-of-the-patients}{%
\subsection{Part 1: Exploring the demographics of the
patients}\label{part-1-exploring-the-demographics-of-the-patients}}

@@

\includegraphics{Final-Submission_files/figure-latex/unnamed-chunk-5-1.pdf}

@@

\includegraphics{Final-Submission_files/figure-latex/unnamed-chunk-6-1.pdf}

@@

\hypertarget{part-2-looking-at-the-data-itself}{%
\subsection{Part 2: Looking at the data
itself}\label{part-2-looking-at-the-data-itself}}

@@

An observation is that many of the variables have 106 NAs. This
indicates that a good fraction {[}106 patients{]} could have been
equally tracked and measured in a less extensive way.

@@

\includegraphics{Final-Submission_files/figure-latex/unnamed-chunk-8-1.pdf}

@@

\hypertarget{part-3-exploring-some-interactions-of-variables}{%
\subsection{Part 3: Exploring some interactions of
variables}\label{part-3-exploring-some-interactions-of-variables}}

@@

\includegraphics{Final-Submission_files/figure-latex/unnamed-chunk-9-1.pdf}

@@

\includegraphics{Final-Submission_files/figure-latex/unnamed-chunk-10-1.pdf}

@@

@@

\includegraphics{Final-Submission_files/figure-latex/unnamed-chunk-11-1.pdf}

@@

\includegraphics{Final-Submission_files/figure-latex/unnamed-chunk-12-1.pdf}

@@

\hypertarget{part-i-classification-algorithm-for-the-status-of-the-patient-using-cart}{%
\section{Part I: Classification Algorithm for the Status of the Patient
using
CART}\label{part-i-classification-algorithm-for-the-status-of-the-patient-using-cart}}

@@

In this report, we have decided to use \{rpart\} because it provides
more flexibility for surrogate splits and the trees are a bit easier to
make attractive looking.

@@

\includegraphics{Final-Submission_files/figure-latex/unnamed-chunk-16-1.pdf}

To gain a further understanding of the data, we can plot a tree yielding
Collection Node style trees. This can help us understand how the data is
split.

@@

\includegraphics{Final-Submission_files/figure-latex/unnamed-chunk-17-1.pdf}

The tree shows us the splits that were done on Age, Bilirubin and
Prothrombin. Interestingly, Stage did not contribute in the tree.

@@

\includegraphics{Final-Submission_files/figure-latex/unnamed-chunk-19-1.pdf}

From the graph we can see that a cp of 0.039 is ideal as it is under the
horizontal (dotted) reference line. We can prune the tree with this CP
value.

@@

\includegraphics{Final-Submission_files/figure-latex/unnamed-chunk-21-1.pdf}

We ca see the pruned tree has cut out some leaf nodes. This would help
in avoiding overfitting the model.

@@

\FunctionTok{kable}\NormalTok{(}
\NormalTok{  conf\_matrix,}
  \AttributeTok{col.names =} \FunctionTok{c}\NormalTok{(}\StringTok{"Prediction/Supervision"}\NormalTok{, }\StringTok{"0"}\NormalTok{, }\StringTok{"1"}\NormalTok{),}
  \AttributeTok{digits =} \DecValTok{3}\NormalTok{,}
  \AttributeTok{booktabs =} \ConstantTok{TRUE}\NormalTok{,}
  \AttributeTok{caption =} \StringTok{"Model 1: Confusion Matrix (0=Deceased, 1=Censored)"}\NormalTok{,}
  \AttributeTok{align =} \StringTok{"c"}
\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
\FunctionTok{kable\_styling}\NormalTok{(}\AttributeTok{latex\_options =} \StringTok{"HOLD\_position"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\caption{\label{tab:unnamed-chunk-23}Model 1: Confusion Matrix (0=Deceased, 1=Censored)}
\centering
\begin{tabular}[t]{lcc}
\toprule
Prediction/Supervision & 0 & 1\\
\midrule
0 & 19 & 9\\
1 & 14 & 43\\
\bottomrule
\end{tabular}
\end{table}

@@

With this, we can also calculate the model's metrics on the test data.

@@

As you can see the model shows good accuracy with 72.9\%. However, the
specificity is an issue with 57.6\%. Now let us compare this with a
different type of model: logistic regression.

@@

\caption{\label{tab:unnamed-chunk-30}Model 1 Confusion Matrix}
\centering
\begin{tabular}[t]{ccc}
\toprule
Predicted/Actual & Censored & Deceased\\
\midrule
Censored & 117 & 68\\
Deceased & 15 & 20\\
\bottomrule
\end{tabular}
\end{table}

@@

This is the role of inference in evaluating our model. The most notable
predictors are Bilirubin, Age, Alk\_Phos, and Prothrombin, each showing
a statistically significant relationship (p \textless{} 0.05) with the
dependent variable. The Intercept and EdemaY have extremely significant
p-values, but the practical significance of EdemaY is questionable due
to its large standard error. Other variables like Spiders, SGOT,
DrugPlacebo, and Copper, while contributing to the model, do not reach
conventional levels of statistical significance (p \textless{} 0.05).

@@

\includegraphics{Final-Submission_files/figure-latex/unnamed-chunk-32-1.pdf}

This figure shows us the Tukey-Anscombe plot using Pearson residuals for
Model 2. In an ideal fit, the residuals should be evenly distributed
about zero with constant mean and variance. The shape of the line
suggests that the model is not capturing someunder lying structure in
the datain extreme cases.

@@

\caption{\label{tab:unnamed-chunk-34}Confusion matrix for Model 2}
\centering
\begin{tabular}[t]{l|r|r}
\hline
  & Censored & Deceased\\
\hline
Censored & 115 & 24\\
\hline
Deceased & 17 & 64\\
\hline
\end{tabular}
\end{table}

@@

\includegraphics{Final-Submission_files/figure-latex/unnamed-chunk-36-1.pdf}

@@

\includegraphics{Final-Submission_files/figure-latex/unnamed-chunk-37-1.pdf}

@@

\includegraphics{Final-Submission_files/figure-latex/unnamed-chunk-39-1.pdf}

@@

Model 1:

\begin{itemize}
\item
  Its ROC curve is above the line of no discrimination, indicating that
  the model has some predictive capabilities
\item
  The AUC is 0.638, which is better than random guessing but suggests
  there's room for improvement.
\end{itemize}

Model 2:

\begin{itemize}
\item
  The ROC curve for Model 2 is significantly above that of Model 1, and
  much closer to the top-left corner, indicating better predictive
  performance.
\item
  The AUC is 0.899, which suggests a good classification performance,
  and it's notably better than Model 1.
\item
  Its ability to discriminate between positive and negative classes is
  superior to that of Model 1.
\end{itemize}

@@

\includegraphics{Final-Submission_files/figure-latex/unnamed-chunk-40-1.pdf}

@@

\caption{\label{tab:unnamed-chunk-42}Confusion Matrix for Test data}
\centering
\begin{tabular}[t]{l|r|r}
\hline
Predicted/Actual & Censored & Deceased\\
\hline
Censored & 30 & 13\\
\hline
Deceased & 3 & 10\\
\hline
\end{tabular}
\end{table}

@@

\includegraphics{Final-Submission_files/figure-latex/unnamed-chunk-43-1.pdf}

\#Comparison between models

Now let us compare the performance of our logistic regression model with
a decision tree model. We will use the same train test split as before.

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#do the ml, take about observational stages vs now quantify, look into the different variables and what they mean.}
\NormalTok{cirrhosisCluster }\OtherTok{\textless{}{-}}\NormalTok{ cirrhosis}
\NormalTok{cirrhosisCluster2 }\OtherTok{\textless{}{-}} \FunctionTok{na.omit}\NormalTok{(cirrhosis)}

@@

\includegraphics{Final-Submission_files/figure-latex/unnamed-chunk-48-1.pdf}

@@

\includegraphics{Final-Submission_files/figure-latex/unnamed-chunk-49-1.pdf}

@@

\includegraphics{Final-Submission_files/figure-latex/unnamed-chunk-51-1.pdf}

@@

\includegraphics{Final-Submission_files/figure-latex/unnamed-chunk-52-1.pdf}

@@

\includegraphics{Final-Submission_files/figure-latex/unnamed-chunk-53-1.pdf}

Now we can go back to our initial goal: comparing how the custers
compare to the Stage variable

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{cirrhosisCluster2}\SpecialCharTok{$}\NormalTok{cluster }\OtherTok{\textless{}{-}}\NormalTok{ hybridCirrhosis}\SpecialCharTok{$}\NormalTok{cluster}
\CommentTok{\# Calculate mean (or median, etc.) for each variable in each cluster}
\FunctionTok{library}\NormalTok{(dplyr)}
\NormalTok{cluster\_summary }\OtherTok{\textless{}{-}}\NormalTok{ cirrhosisCluster2 }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{group\_by}\NormalTok{(cluster) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{summarise\_all}\NormalTok{(}\FunctionTok{funs}\NormalTok{(}\FunctionTok{mean}\NormalTok{(., }\AttributeTok{na.rm =} \ConstantTok{TRUE}\NormalTok{))) }\CommentTok{\# Replace mean with median or any other function as necessary}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning: `funs()` was deprecated in dplyr 0.8.0.
## i Please use a list of either functions or lambdas:
## 
## # Simple named list: list(mean = mean, median = median)
## 
## # Auto named with `tibble::lst()`: tibble::lst(mean, median)
## 
## # Using lambdas list(~ mean(., trim = .2), ~ median(., na.rm = TRUE))
## Call `lifecycle::last_lifecycle_warnings()` to see where this warning was
## generated.
\end{verbatim}

\begin{verbatim}
## Warning: There were 20 warnings in `summarise()`.
## The first warning was:
## i In argument: `Status = mean(Status, na.rm = TRUE)`.
## i In group 1: `cluster = 1`.
## Caused by warning in `mean.default()`:
## ! argument is not numeric or logical: returning NA
## i Run `dplyr::last_dplyr_warnings()` to see the 19 remaining warnings.
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{kable}\NormalTok{(cluster\_summary }\SpecialCharTok{\%\textgreater{}\%}\NormalTok{ dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(}\SpecialCharTok{{-}}\NormalTok{Status, }\SpecialCharTok{{-}}\NormalTok{Drug, }\SpecialCharTok{{-}}\NormalTok{ Sex, }\SpecialCharTok{{-}}\NormalTok{Edema, }\SpecialCharTok{{-}}\NormalTok{Stage), }\AttributeTok{label =} \StringTok{"Cluster Summary"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{tabular}{r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r}
\hline
cluster & N\_Days & Age & Ascites & Hepatomegaly & Spiders & Bilirubin & Cholesterol & Albumin & Copper & Alk\_Phos & SGOT & Tryglicerides & Platelets & Prothrombin & Age\_Years\\
\hline
1 & 1475.862 & 23503.95 & 0.1724138 & 0.5517241 & 0.3103448 & 4.034483 & 327.0862 & 3.337069 & 107.77586 & 1492.241 & 120.0178 & 126.7414 & 248.3276 & 11.15517 & 64.35030\\
\hline
2 & 2579.950 & 18505.65 & 0.1000000 & 0.6500000 & 0.3500000 & 4.275000 & 376.3500 & 3.372000 & 149.15000 & 8394.200 & 136.8590 & 145.4000 & 263.0500 & 10.98500 & 50.66571\\
\hline
3 & 2074.659 & 13933.03 & 0.0227273 & 0.4204545 & 0.2954545 & 2.842045 & 426.8977 & 3.629773 & 93.68182 & 1582.877 & 130.9384 & 119.4205 & 281.6591 & 10.46591 & 38.14657\\
\hline
4 & 2058.918 & 18734.59 & 0.0454545 & 0.5454545 & 0.2636364 & 3.186364 & 349.1182 & 3.547546 & 93.94545 & 1430.342 & 118.5102 & 124.7818 & 252.7182 & 10.68455 & 51.29251\\
\hline
\end{tabular}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{ggplot}\NormalTok{(cirrhosisCluster2, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =} \FunctionTok{factor}\NormalTok{(cluster), }\AttributeTok{fill =} \FunctionTok{factor}\NormalTok{(Stage))) }\SpecialCharTok{+}
  \FunctionTok{geom\_bar}\NormalTok{(}\AttributeTok{position =} \StringTok{"dodge"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}\AttributeTok{title =} \StringTok{"Distribution of Stages within Clusters"}\NormalTok{,}
       \AttributeTok{x =} \StringTok{"Cluster"}\NormalTok{,}
       \AttributeTok{y =} \StringTok{"Count"}\NormalTok{,}
       \AttributeTok{fill =} \StringTok{"Stage"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_minimal}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{Final-Submission_files/figure-latex/unnamed-chunk-54-1.pdf}

We can see that the clusters are not evenly distributed. This is a good
sign, as it means that the clusters are not just a random grouping of
the data.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\hypertarget{author-contributions}{%
\section{Author Contributions}\label{author-contributions}}

\hypertarget{referencescitation-style-is-your-choice-but-all-sources-should-be-documented-both-in-text-and-in-the-references-section.-this-includes-where-you-got-your-data.}{%
\section{References--Citation style is your choice, but all sources
should be documented (both in text and in the References section). This
includes where you got your
data.}\label{referencescitation-style-is-your-choice-but-all-sources-should-be-documented-both-in-text-and-in-the-references-section.-this-includes-where-you-got-your-data.}}

\hypertarget{code-appendix}{%
\section{Code Appendix}\label{code-appendix}}
